---
title: "data_wrangling"
author: "Bin Yang"
date: "11/11/2020"
output: html_document
---

```{r setup}
library(tidyverse)
library(tidymodels)
library(plotly)
library(viridis)
library(modelr)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

#### EDA
Summary of the dataset:  

```{r}
hotel_df = 
  read.csv("./data/hotel_bookings.csv",na = c("", "NA", "NULL")) %>% 
  mutate_if(is.character, as.factor)
```
  

  
produce summary tables for categorical variables:  

```{r}
# skimr::skim(hotel_df)
# hotel_df %>% 
#   select(which(sapply(.,is.character))) %>% 
#   gtsummary::tbl_summary(by = hotel)
```

Possible plotly plots! (will edit labels etc. later)

Booking and cancellation by month:   I think color is pretty awful rn......

Bar plot:  
```{r}
# hotel_df %>% 
#   select(arrival_date_month, is_canceled) %>% 
#   mutate(arrival_date_month = as.factor(arrival_date_month),
#          is_canceled = as.factor(is_canceled)) %>% 
#   count(arrival_date_month, is_canceled) %>% 
#   mutate(arrival_date_month = fct_reorder(arrival_date_month, n)) %>% 
#   plot_ly(x = ~arrival_date_month, y = ~ n, color = ~is_canceled, type ="bar", colors = "viridis")
#   
```

Does the hotel charged differently for different customer type?   

need to do: modify axies, deal with outliers! 

boxplot:   
```{r}
# hotel_df %>% 
#   select(hotel, customer_type, adr) %>% 
#   mutate(hotel= as.factor(hotel), 
#          customer_type = as.factor(customer_type),
#          customer_type = fct_reorder(customer_type, adr))%>%
#   plot_ly(y = ~adr, x = ~customer_type, color = ~hotel, type ="box" , colors = "viridis") %>% 
#   layout(boxmode = "group", yaxis = list(range=c(0,800)))
```

Home country of guests! 

a lot to improve... hover etc...
choropleth maps:  
```{r}
# hotel_df %>% 
#   select(country, is_canceled) %>% 
#   filter(is_canceled == 0) %>% 
#   group_by(country) %>% 
#   summarise(
#     total_booking = n(),
#   ) %>% 
#   mutate(
#     booking_percent = round(total_booking/sum(total_booking)*100, 2)
#   ) %>% 
#   plot_ly(locations = ~country, z = ~booking_percent, type ="choropleth",
#           colorscale="Viridis", zmin = 0, zmax = 30)
```

#### Logistic Regression :    

data preprocessing:  

total NAs:  
```{r}
# table_NA=
#   hotel_df %>% 
#   summarise_all(funs(sum(is.na(.))))
# 
# t(table_NA) %>% 
#   knitr::kable()
```

```{r}
# hotel_ml = 
#   hotel_df %>% 
#   mutate(
#     arrival_season = case_when(
#       arrival_date_month %in% c("October","November","December") ~ "Fall",
#       arrival_date_month %in% c("January","February","March") ~ "Winter",
#       arrival_date_month %in% c("April","May","June") ~ "Spring",
#       TRUE ~ "Summer"
#     ),
#     arrival_season = as.factor(arrival_season),
#     reserved_room_type = as.character(reserved_room_type),
#     assigned_room_type = as.character(assigned_room_type),
#     false_room_assignment = case_when(
#       reserved_room_type == assigned_room_type ~ "0",
#       reserved_room_type != assigned_room_type ~ "1"
#     ),
#     false_room_assignment = as.factor(false_room_assignment),
#     is_canceled = as.factor(is_canceled)) %>% 
#     select(-c(arrival_date_year, arrival_date_week_number, arrival_date_day_of_month,
#               arrival_date_month, reservation_status_date, reservation_status, 
#               agent, company, country, reserved_room_type, assigned_room_type)) 
```

read in dataset:
```{r}
hotel_ml = 
  read_csv("./data/hotel_ml.csv") %>% 
  mutate(
    is_canceled = as.factor(is_canceled)
  ) %>% 
  mutate_if(is.character, as.factor) %>% 
  select(-arrival_date_month) %>% 
  drop_na() 
```

```{r}
table_NA=
  hotel_ml %>%
  summarise_all(funs(sum(is.na(.))))

t(table_NA) %>%
  knitr::kable()
```

We used stepwise selection and concluded that all of our existing variables have statistical significance.
```{r}
fullmod = glm(is_canceled ~. , data = hotel_ml, family = binomial())
finalmod = stats::step(fullmod, direction = "both")
```
```

train/test split:
```{r}
set.seed(1234)
train_df = sample_frac(hotel_ml, 0.8)
test_df = anti_join(hotel_ml, train_df)
```

```{r}
logistic_fit =  
  train_df %>% 
  glm(is_canceled ~. , data = ., family = binomial())
```

Model output: 
```{r}
logistic_fit %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate)) %>%
  select(term, log_OR = estimate, OR, p.value) %>% 
  arrange(desc(OR)) %>% 
  knitr::kable(digits = 3, caption = "Logistic Regression Output") 
```

**Comment** As we can see from the output, deposit type: non refund has the largest odds ratio(173.389), indicating that the guests with non refundable test are significantly more likely to cancel the bookings compared to no deposit guests. This finding is rather counter-intuitive and also raises the question about the validity of the dataset.  

We furthur generated the most important variables of our model:  
```{r variable importance}
caret::varImp(logistic_fit) %>%
 arrange(desc(Overall)) %>%
 knitr::kable(digits = 3)
```

We will look more into the four most important features:  
* Total special requests
* deposit type 
* adr  
(can i use plots from ead?)

Since we are questioning the validity of deposit type, we fitted the logistic regression models withou the deposit type. We then compared the prediction accuracy of the two models using 10 fold cross validation: 
```{r}
cv_df = crossv_mc(hotel_ml, 10) 
cv_df = 

```

```{r}
# Make predictions
probabilities <- logistic_fit%>% predict(test_df, type = "response")
predicted.classes <- ifelse(probabilities > 0.5, "1", "0")
# Prediction accuracy
observed.classes <- test_df$is_canceled
mean(predicted.classes == observed.classes, na.rm = T)
```


use cross validation
```{r}
# cv_df = crossv_mc(hotel_ml, 10)
# cv = 
# cv_df %>% 
#   mutate(
#     train = map(train, as_tibble),
#     test = map(test, as_tibble),
#     fit_logistic = map(train, ~glm(is_canceled ~. , data = .x, family = binomial()))
#   )
#
```




fit logistic regression: 
```{r}
# fit_logistic = 
#   train_df %>% 
#   glm(is_canceled ~. , data = ., family = binomial())
```


model evaluation & diagnoistics 
possible outliers:


Validation of Predicted Values
```{r}
# test_predicted <- predict(fit_logistic, newdata = test_df, type = "response")
```



