---
title: "data_wrangling"
author: "Bin Yang"
date: "11/11/2020"
output: html_document
---

```{r setup}
library(tidyverse)
library(plotly)
library(viridis)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

#### EDA
Summary of the dataset:  

```{r}
hotel_df = read.csv("./data/hotel_bookings.csv")
```
  
produce summary tables for categorical variables:  

```{r}
# skimr::skim(hotel_df)
# hotel_df %>% 
#   select(which(sapply(.,is.character))) %>% 
#   gtsummary::tbl_summary(by = hotel)
```

Possible plotly plots! (will edit labels etc. later)

Booking and cancellation by month:   I think color is pretty awful rn......

Bar plot:  
```{r}
hotel_df %>% 
  select(arrival_date_month, is_canceled) %>% 
  mutate(arrival_date_month = as.factor(arrival_date_month),
         is_canceled = as.factor(is_canceled)) %>% 
  count(arrival_date_month, is_canceled) %>% 
  mutate(arrival_date_month = fct_reorder(arrival_date_month, n)) %>% 
  plot_ly(x = ~arrival_date_month, y = ~ n, color = ~is_canceled, type ="bar", colors = "viridis")
  
```

Does the hotel charged differently for different customer type?   

need to do: modify axies, deal with outliers! 

boxplot:   
```{r}
hotel_df %>% 
  select(hotel, customer_type, adr) %>% 
  mutate(hotel= as.factor(hotel), 
         customer_type = as.factor(customer_type),
         customer_type = fct_reorder(customer_type, adr))%>%
  plot_ly(y = ~adr, x = ~customer_type, color = ~hotel, type ="box" , colors = "viridis") %>% 
  layout(boxmode = "group", yaxis = list(range=c(0,800)))
```

Home country of guests! 

a lot to improve... hover etc...
choropleth maps:  
```{r}
hotel_df %>% 
  select(country, is_canceled) %>% 
  filter(is_canceled == 0) %>% 
  group_by(country) %>% 
  summarise(
    total_booking = n(),
  ) %>% 
  mutate(
    booking_percent = round(total_booking/sum(total_booking)*100, 2)
  ) %>% 
  plot_ly(locations = ~country, z = ~booking_percent, type ="choropleth",
          colorscale="Viridis", zmin = 0, zmax = 30)
```

#### Modeling:  

total NAs:  
```{r}
table_NA=
  hotel_df %>% 
  summarise_all(funs(sum(is.na(.))))

t(table_NA) %>% 
  knitr::kable()
```


data preprocessing:  
```{r}
hotel_df = 
  hotel_df %>% 
  mutate_if(is.character, as.factor) %>% 
  drop_na() %>% 
  mutate(id =  row_number())
```


train/test split:  
```{r}
train_df = 
  sample_n(hotel_df, floor(0.8*nrow(hotel_df)))

test_df =
  anti_join(hotel_df, train_df, by = "id")  
```

proportion of cancellation of train/test dataset:  
```{r}
train_df %>% 
  group_by(is_canceled) %>% 
  summarise(
    number_of_observations =   n(),
    proportions = round(number_of_observations / nrow(train_df), 2)
  ) %>% 
  knitr::kable()
  
test_df %>% 
  group_by(is_canceled) %>% 
  summarise(
    number_of_observations =   n(),
    proportions = round(number_of_observations / nrow(test_df), 2)
  ) %>% 
  knitr::kable()
```




correlation:  
```{r}
library(`ggcorrplot`)
```






