---
title: "data_wrangling"
author: "Bin Yang"
date: "11/11/2020"
output: html_document
---

```{r setup}
library(tidyverse)
library(tidymodels)
library(plotly)
library(viridis)
library(modelr)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

#### EDA
Summary of the dataset:  

```{r}
hotel_df = 
  read.csv("./data/hotel_bookings.csv",na = c("", "NA", "NULL")) %>% 
  mutate_if(is.character, as.factor)
```
  

  
produce summary tables for categorical variables:  

```{r}
# skimr::skim(hotel_df)
# hotel_df %>% 
#   select(which(sapply(.,is.character))) %>% 
#   gtsummary::tbl_summary(by = hotel)
```

Possible plotly plots! (will edit labels etc. later)

Booking and cancellation by month:   I think color is pretty awful rn......

Bar plot:  
```{r}
# hotel_df %>% 
#   select(arrival_date_month, is_canceled) %>% 
#   mutate(arrival_date_month = as.factor(arrival_date_month),
#          is_canceled = as.factor(is_canceled)) %>% 
#   count(arrival_date_month, is_canceled) %>% 
#   mutate(arrival_date_month = fct_reorder(arrival_date_month, n)) %>% 
#   plot_ly(x = ~arrival_date_month, y = ~ n, color = ~is_canceled, type ="bar", colors = "viridis")
#   
```

Does the hotel charged differently for different customer type?   

need to do: modify axies, deal with outliers! 

boxplot:   
```{r}
# hotel_df %>% 
#   select(hotel, customer_type, adr) %>% 
#   mutate(hotel= as.factor(hotel), 
#          customer_type = as.factor(customer_type),
#          customer_type = fct_reorder(customer_type, adr))%>%
#   plot_ly(y = ~adr, x = ~customer_type, color = ~hotel, type ="box" , colors = "viridis") %>% 
#   layout(boxmode = "group", yaxis = list(range=c(0,800)))
```

Home country of guests! 

a lot to improve... hover etc...
choropleth maps:  
```{r}
# hotel_df %>% 
#   select(country, is_canceled) %>% 
#   filter(is_canceled == 0) %>% 
#   group_by(country) %>% 
#   summarise(
#     total_booking = n(),
#   ) %>% 
#   mutate(
#     booking_percent = round(total_booking/sum(total_booking)*100, 2)
#   ) %>% 
#   plot_ly(locations = ~country, z = ~booking_percent, type ="choropleth",
#           colorscale="Viridis", zmin = 0, zmax = 30)
```


```{r data,echo=FALSE, message=F}
hotel_whole_df <- read.csv("./data/hotel_bookings.csv",na = c("", "NA", "NULL")) 
hotel_df <- hotel_whole_df %>% filter(arrival_date_year==2016)
```

```{r}
hotel_df %>% 
  select(is_canceled, adr) %>% 
  group_by(is_canceled, adr) %>% 
  summarize(
    count = n()
  ) %>% 
  filter(count > 10, is_canceled == 1) %>% 
  drop_na() %>% 
  mutate(
    fv = lm(count ~ adr, .) %>%  fitted.values()
  ) %>% 
  plot_ly(x = ~adr, y = ~count, mode = "markers") %>%
  add_markers(y = ~count) %>% 
  add_trace(x = ~adr, y = ~fv, mode = "lines") %>% 
  layout(title = "Average Daily Rate (ADR) and Cancellation", 
         xaxis = list(title = "ADR"), 
         yaxis = list(title = "Cancellation Count"),
         showlegend = F)
```

