---
title: "Project Report"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true

---

```{r setup, include=FALSE, echo=FALSE, message=F}
library(dplyr)
library(tidyverse)
library(plotly)
library(ggplot2)
library(modelr)
library(ROCR)
library(pROC)
library(purrr)
library(readr)
library(tidyr)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Motivation 
## Related work
* Antonio N., Almeida A., Nunes L. (2019). Hotel Booking Demand Datasets. Data in Brief. Vol 22. 41-49.

## Initial Questions
* Where do the guests come from?   
* How does price vary over the year?    
* What are the potential factors which could influence the cancellation of both hotels?

## Data
```{r data,echo=FALSE, message=F}
hotel_whole_df <- read.csv("./data/hotel_bookings.csv",na = c("", "NA", "NULL")) 
hotel_df <- hotel_whole_df %>% filter(arrival_date_year==2016)
```

> **Primary Data Source** 

 [Hotel Booking Demand Dataset](https://www.sciencedirect.com/science/article/pii/S2352340918315191) 
 
> **Variable of Interest**

The original dataset contains information from 2015 to 2017. To avoid the caveat of time series and for a speedier performance of our model, we only included the year 2016 for our analysis. In order to identify any seasonality trend over the year, we created a new variable containing the arrival season of the guests. We then mutated the character variables as factors and additionally some numeric variables that would be better explained as factors. To avoid collinearlity, we kept arrival season and exluced arrival month. Finally, to satisfy our aim of predicting booking cancelation, we selected the following variables that we desired to include in our model:  

  1. Outcome:\
 _IsCanceled_: Values indicating if the booking was canceled (1) or not (0).\
 \
  2. Potential Predictors:\
  * _Categorical Variables_:\
    + _ArrivalDateMonth_: Month of arrival date with 12 categories: “January” to “December”.
    + _CustomerType_: Types of booking, assuming one of the four categories:\
-Contract:  when the booking has an allotment or other type of contract associated with it (long-term stay);\
-Group: when the booking is associated with a group;\
-Transient: when the booking is not part of a group or contract and is not associated with other transient bookings (these are guests who are predominantly on-the-move and seek short (and often urgent) hotel-stays)\
-Transient-party:  when the booking is transient but is associated with at least another transient booking.\
    + _DepositType_: Indication on if the customer made a deposit to guarantee the booking. The variable can assume 3 categories:
-No Deposit;\
-Non Refund: Deposit value equals or exceeds the total stay cost;\
-Refundable: Deposit value under the total stay cost.\
    + _DistributionChannel_: Booking distribution channel (“TA” means travel agent; “TO” means tour operator).\
    + Hotel: City Hotel or Resort Hotel. \
    + _MarketSegment_: Market segment designation (“TA” means travel agent; “TO” means tour operator). \
    + _Meal_: Type of meal booked.\
-Undefined/SC: no meal package;\
-BB: Bed & breakfast;\
-HB: Half board (breakfast and one other meal)\
-FB: Full board (breakfast, lunch, and dinner)\
    + _ReservedRoomType_: Code of room type reserved. Code is presented instead of designation for anonymity reasons.\
    
  * _Numeric Variables_:\
    + _ADR_: Average daily rate.\
    + _BookingChanges_: Number of changes/amendments made to the booking from the moment the booking was entered on the PMS until the moment of check-in or cancellation. \
    + _PreviousCancellations_: Number of previous bookings that were canceled by the customer prior to the current booking.\
    + _LeadTime_: Number of days that elapsed between the entering date of the booking into the PMS and the arrival date. \
    + _TotalOfSpecialRequests_: Number of special requests made by the customer (e.g. twin bed or high floor)\


> **Data Cleaning and Preprocessing**

  + Convert missing value and NAs (both capital and lower-case) to na for R recognition.
  + Convert numeric to factor variables if needed.
  + Create _ArrivalSeason_ variable based on _ArrivalDateMonth_ to investigate the effect of seasonality on hotel booking demands.
  + Create _FalseRoomAssignment_ variable by comparing the difference between _ReservedRoomType_ and _AssignmedRoomType_. 
  




## Exploratory Analysis
We first used a choropleth map and a donut chart to visualize the proportions of guests' nationality. We observed that most guests come from Europe and top three nationality are Portuguese, with British and French guest come second and third.     

We then tried to explore how the ADR(average daily rate) change over time. Therefore, we used group by summary to produce the mean ADR by month, and plotted these values using line chart. To illustrate how the ADR varies, we also added confidence band to the average line. We observed that the ADR is pretty stable across the years for city hotel, which is reasonable according to the booking demand pattern. For resort hotel, we see a sharp increase in ADR from May to November with the price level peaking during summer. This finding is also intuitive since we expect the demand for resort hotel increases in summer. 

Furthermore, we want to explore the potential factors which can influence the cancellation of both hotels. Therefore, we used the proportion test to produce the proportion of cancellation with each of our variables of interest to observe some interesting changes. 

First of all, we investigated the proportion of cancellation with assigned room type. For the raw data set, there is also a variable called 'reserved room type'. Our consideration is that this room type was chosen by guests themselves, therefore, the proportion of cancellation with this variable might not plausible enough to tell any underlying effects. However, the assigned room type was not determined by guests, so the proportion of cancellation with this variable might give us any clues of the initial relationship about these two variables. Firstly, we used group by summary to produce the number of each assigned room type for both hotels to see the distribution of assigned room type. For both hotels, the number of assigned room type P is extremely low compared to the size of other assigned room type. Therefore, we could ignore the effect brought by this room type on the cancellation proportion. Then, we used group by summary again to produce the total number of assigned guests and the number of assigned guests who have canceled their order. Next, prop_test and map2 function were applied to obtain the proportion of the cancellation corresponding to each assigned room type. These results were then plotted as a bar graph to virtually show the cancellation proportion of each assigned room type by hotel. Indeed, the plot indicated that the cancellation proportion for assigned room type P is 1 for both hotels, which is a non-comparable extreme value due to its non-comparable sample size mentioned before. Therefore, we turned to look at other proportion values. Among the resting results, we found that the cancellation proportion is distinctly high among guests who were assigned room type A, G, H with comparable sizes for Resort Hotel. As for City Hotel, the cancellation proportion is found to be highest among guests who were assigned room type A with comparable size. Based on these observations, we could infer that the room types mentioned above might not meet the expectations and needs of guests compared to their originally reserved room type so that make them cancel the order.

Secondly, we investigated the proportion of cancellation with distribution channel. The same process as above was applied to these two variables to generate the distribution and proportion plots. For the distribution of the booking channels, we found that most of the guests for both hotels chose to book the room through TA/TO, which are “Travel Agents” and “Tour Operators” respectively. Then, from the proportion plot, we found that the cancellation proportion among this group was also the highest compared to that of the other groups. Therefore, we could infer that there might exit some unpleasant or disappointed experiences for guests by using TA/TO channels to book the hotel.

Thirdly, we investigated the proportion of cancellation with meal type. The same process as above was applied to these two variables to generate the distribution and proportion plots. For the distribution of the booked meal types, we found that the size of the booked FB meal for both hotels and SC meal for Resort hotel were extremely small compared to that of the other booked meal types. Therefore, considering the result of these meal types might not be comparable to others, we just ignored their effects on the cancellation proportion. Then, according to the proportion plot, we've found that the cancellation proportion of FB meal was indeed distinctly high for Resort Hotel, which met our previous consideration so that we ignored this effect. Otherwise, among all other booked meal types, for Resort Hotel, we found that the cancellation proportion was highest among guests who booked HB meal with comparable size; for City Hotel, the cancellation proportion is highest among guests who booked BB meal. Therefore, we could infer that these two meal types might not meet the expectations and needs of guests, which leads them to cancel the order.

Fourthly, we investigated the proportion of cancellation with repeated guest. The same process as above was applied to these two variables to generate the distribution and proportion plots. From the distribution plot, we found that the main guest source was from non-repeated guest for both hotels. Then from the proportion plot, we found that for both hotels, the cancellation proportion is significantly higher among non-repeated guests, which met the common expectation for the repeated-guest effect. 



## Statistical Analysis & Prediction Model


### Overview
Using existing data, we investigated factors that are associated with - and can potentially be used to predict - hotel booking cancellation.\
Two types of models are used for prediction - logistic regression and random forest. \
Categorical variables are selected based on our hypothesis on its association with cancellation, and for numeric variables, we make the selected based on AUC scores.\
Overall, the logistic regression model achieves an accuracy of 75%. The accuracy for the random forest model is about 86%.


### Variable Selection
The dataset is quite large (119,390 rows, 32 variables). For a speedy performance of our models, we first selected instances in year 2016 alone (the dataset ranges from 2015 to 2017), which resulted in 56,707 rows.\
\
For categorical variables, we selected 8 out of 14 variables based on our hypothesis on their association with _IsCanceled_.\
\
For numeric variables, we selected 5 variables based on their AUC scores.
```{r AUC,echo=FALSE, message=F }

 hotel_df_xyl <- read.csv("./data/hotel_bookings.csv",na = c("", "NA", "NULL")) %>%
   select(where(is.numeric)) %>% select(-agent)

 variable_list <- colnames(hotel_df_xyl)
 variable_list <- variable_list[-1]

 compute_auc <- function(x) {
   as.numeric(auc(roc(hotel_df_xyl[["is_canceled"]][!is.na(hotel_df_xyl[[x]])], hotel_df_xyl[[x]][!is.na(hotel_df_xyl[[x]])])))
 }

 auc_result<-as_tibble(t(combn(variable_list, 1)),
           .name_repair = ~ c("var_name"))%>%
   mutate(
     auc = map_dbl(.x = var_name, compute_auc)
   ) %>%
   arrange(-auc)
 auc_result %>% knitr::kable(digits = 3)

```

An AUC score equals 0.5 implies that the variable is not discriminable between two classes. Thus, we chose variables that had the greatest distances to 0.5.\
From the AUC score table above, we chose _LeadTime_, _TotalOfSpecialRequests_, _PreviousCancellations_, _adr_, and _BookingChanges_ for our models.



### Logistic Regression  
```{r data import,echo=FALSE, message=F}
hotel_ml = 
  read_csv("./data/hotel_ml.csv") %>% 
  mutate(
    is_canceled = as.factor(is_canceled)
  ) %>% 
  mutate_if(is.character, as.factor) %>% 
  select(-arrival_date_month) %>% 
  drop_na() 
```

```{r, include = FALSE, Fecho=FALSE, message=F}
fullmod = glm(is_canceled ~. , data = hotel_ml, family = binomial())
finalmod = 
  stats::step(fullmod, direction = "both") %>% 
  broom::tidy() %>% 
  knitr::kable()
```
We first tried to predict the output using Logistic regression. The first model we used included all the possible predictors we selected. To perform variable selection, we used stepwise regression. The result shows that the all of our existing variables have statistical significance. We thus fitted this model to our dataset, and generated the following output for the variables with significant odds ratios:  
```{r test/train split,echo=FALSE, message=F, warning=FALSE}
set.seed(1234)
train_df = sample_frac(hotel_ml, 0.8)
test_df = anti_join(hotel_ml, train_df)

logistic_fit =  
  train_df %>% 
  glm(is_canceled ~. , data = ., family = binomial())

logistic_fit %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate)) %>%
  select(term, log_OR = estimate, OR, p.value) %>% 
  filter(OR > 1.9 | OR < 0.2) %>% 
  arrange(desc(OR)) %>% 
  knitr::kable(digits = 3, caption = "Logistic Regression Output") 
```
**Comment** As we can see from the output, deposit type: non refund has the largest odds ratio(173.389), indicating that the guests with non refundable test are significantly more likely to cancel the bookings compared to no deposit guests. This finding is rather counter-intuitive and also raises the question about the validity of the dataset.  
<br>

#### Variable Importance 
We furthur generated the most important variables of our model:  
```{r variable importance,echo=FALSE, message=F}
caret::varImp(logistic_fit) %>%
  arrange(desc(Overall)) %>%
  filter(Overall > 10) %>% 
  knitr::kable(digits = 3, caption = "Variable Importance")
```

##### We will look more into the three most important features:  
* Total special requests
* deposit type 
* lead time
* adr 

<br> 


##### The Special Request Effect

```{r specialrequest, echo=F, message=F, warning=F}
special_request_df <- 
  hotel_df %>%
  mutate(is_canceled = recode(is_canceled, "1" = "Canceled", "0" = "Not Canceled")) %>%
  select(hotel, is_canceled, total_of_special_requests)


special_request_prop <- 
  special_request_df %>%
  group_by(total_of_special_requests, hotel) %>%
  summarize(
    guest_total = n(), 
    guest_canceled = sum(is_canceled == "Canceled")
  ) %>%
  mutate(
    prop_test = map2(.x = guest_canceled, .y = guest_total, ~prop.test(x = .x, n = .y)),
    tidy_test = map(.x = prop_test, ~broom::tidy(.x))
  ) %>%
  select(-prop_test) %>%
  unnest() %>%
  select(total_of_special_requests, hotel, estimate, conf.low, conf.high)


special_request_prop_plot <- 
  special_request_prop %>%
  mutate(text_label = str_c("Number of Special Requests: ", total_of_special_requests, ", Cancellation Proportion: ", estimate)) %>%
    plot_ly(
    x = ~total_of_special_requests, y = ~estimate, color = ~hotel,
    type = "bar", colors = "viridis", text = ~text_label
  ) %>%
  layout(title = "Canceled Proportion with Total Number of Special Request",
         xaxis = list(title = "Number of Special Requests"),
         yaxis = list(title = "Proportion of Cancellation", range = c(0, 1)),
         barmode = "group"
)

special_request_prop_plot
```

Based on the proportion plot, we found that, for city hotel, the cancellation proportion is the highest when the guests didn't ask for any special request. Similarly, for resort hotel, the cancellation proportion is also higher when guests didn't ask for any special request. 
 
<br> 

##### The Deposit Type Effect

```{r deposit, echo=F, message=F, warning=F}
deposit_df <- 
  hotel_df %>%
  mutate(is_canceled = recode(is_canceled, "1" = "Canceled", "0" = "Not Canceled")) %>%
  select(hotel, is_canceled, deposit_type)


deposit_prop <- 
  deposit_df %>%
  group_by(deposit_type, hotel) %>%
  summarize(
    guest_total = n(), 
    guest_canceled = sum(is_canceled == "Canceled")
  ) %>%
  mutate(
    prop_test = map2(.x = guest_canceled, .y = guest_total, ~prop.test(x = .x, n = .y)),
    tidy_test = map(.x = prop_test, ~broom::tidy(.x))
  ) %>%
  select(-prop_test) %>%
  unnest() %>%
  select(deposit_type, hotel, estimate, conf.low, conf.high)


deposit_prop_plot <- 
  deposit_prop %>%
  mutate(text_label = str_c("Deposit Type: ", deposit_type, ", Cancellation Proportion: ", estimate)) %>%
    plot_ly(
    x = ~deposit_type, y = ~estimate, color = ~hotel,
    type = "bar", colors = "viridis", text = ~text_label
  ) %>%
  layout(title = "Canceled Proportion with Deposit Type",
         xaxis = list(title = "Deposit Type"),
         yaxis = list(title = "Proportion of Cancellation", range = c(0, 1)), 
         barmode = "group"
)

deposit_prop_plot
```

Based on the proportion plot, we found that for both resort and city hotels, reservations with non-refundable deposit have about 94% and 99% of cancellation respectively, which is significantly higher than other two types of deposits. For reservations with refundable deposit, we found that the cancellation of city hotel is relatively higher than that of resort hotel. This counter intuitive finding confirmed with the previous odds ratios. We therefore suspected the validity of this variable.  

<br> 

#### The Lead Time Effect

```{r leadtime, echo=F, message=F, warning=F}
leadtime <-   
  hotel_df %>%
  select(hotel, is_canceled, lead_time) %>%
  mutate(is_canceled = recode(is_canceled, "1" = "Canceled", "0" = "Not Canceled"))

leadtime_box <- 
  leadtime %>%
  plot_ly(
    x = ~is_canceled, y = ~lead_time,
    type = "box", color = ~hotel, 
    colors = "viridis"
  ) %>%
  layout(title = "Cancellation Distribution with Lead Time",
         xaxis = list(title = "cancellation"),
         yaxis = list(title = "lead time (days)"),
         boxmode = "group"
         )

leadtime_box
```
Based on the box plot, we found that, for both city and resort hotels, the medians of lead time among being canceled reservations are relatively higher than those of not being canceled. 

<br>

##### The ADR Effect
```{r echo=F, message=F, warning=F}
hotel_df %>% 
  select(is_canceled, adr) %>% 
  group_by(adr) %>% 
  summarize(
    total = n(),
    canceled = sum(is_canceled == 1),
    prop = round(canceled/total, 2)
  ) %>% 
  filter(total > 10) %>% 
  drop_na() %>% 
  mutate(
    fv = lm(prop ~ adr, .) %>%  fitted.values()
  ) %>% 
  mutate(text_label = str_c("Cancellation Proportion: ", prop, ", ADR: ", adr)) %>%
  plot_ly(x = ~adr, y = ~prop, mode = "markers", text = ~text_label) %>%
  add_markers(y = ~prop) %>% 
  add_trace(x = ~adr, y = ~fv, mode = "lines") %>% 
  layout(title = "Average Daily Rate (ADR) and Cancellation", 
         xaxis = list(title = "ADR"), 
         yaxis = list(title = "Cancellation Proportion"),
         showlegend = F)
```

From this plot, there is no clear trend detected. We see that the proportion of cancelation slight increases with increase in ADR.
    
#### Cross Validation  
Since we are questioning the validity of deposit type, we fitted the logistic regression models without the deposit type. We then compared the prediction accuracy of the two models:
```{r ,echo=FALSE, message=F, warning = FALSE}
logistic_fit_2 = 
  train_df %>% 
  glm(is_canceled ~.-deposit_type , data = ., family = binomial())

test_df %>% 
  add_predictions(logistic_fit) %>% 
  mutate(
    prob_model_1 = boot::inv.logit(pred),
    pred_mod1 = 
      case_when(
        prob_model_1 < 0.5 ~ 0,
        prob_model_1 >= 0.5 ~ 1
         )) %>% 
  select(-pred, -prob_model_1) %>% 
  add_predictions(logistic_fit_2) %>% 
   mutate(
    prob_model_2 = boot::inv.logit(pred),
    pred_mod2 = 
      case_when(
        prob_model_2 < 0.5 ~ 0,
        prob_model_2 >= 0.5 ~ 1
         )) %>% 
  select(is_canceled, pred_mod1, pred_mod2) %>% 
  summarise(
    model1_accuracy = sum(is_canceled==pred_mod1)/ nrow(test_df),
    model2_accuracy = sum(is_canceled==pred_mod2)/ nrow(test_df)
  ) %>% 
  knitr::kable(digits = 3, 
               caption = "Model Accuracy",  
               col.names = c("Model1 accuracy", "Model2 accuracy"))
```
We can see that the two models have roughly the same prediction accuracy! We also generated the ROC for comparison:  
```{r echo=FALSE, message=F}
library(ROCR)

test.predicted.m1 <- predict(logistic_fit, newdata = test_df, type = "response")

test.predicted.m2 <- predict(logistic_fit_2, newdata = test_df, type = "response")

par(mfrow=c(1, 2))

prediction(test.predicted.m1, test_df$is_canceled) %>%
  performance(measure = "tpr", x.measure = "fpr") %>%
  plot()

prediction(test.predicted.m2, test_df$is_canceled) %>%
  performance(measure = "tpr", x.measure = "fpr") %>%
  plot()
```

<br>

#### Model Diagnosis 
```{r}
train_df =
  broom::augment(logistic_fit_2) %>% 
  mutate(index = 1:n()) 

ggplot(data = train_df, aes(index, .std.resid, color = "viridis")) + 
  geom_point(alpha = .5) +
  geom_ref_line(h = 3) + 
  labs(title = "Residual Deviance plot")

plot(logistic_fit_2, which = 4, id.n = 5)
```

<br>
We created the residual deviance plot to detect outliers. And as shown in the plot, those standardized residuals that exceed 3 represent possible outliers and may deserve closer attention. Similarly, plotting the cook's distance also signals some potential outliers in our dataset.  

## Random Forest

```{r rf_setup, echo = F, message=F}
library(reshape2)
library(randomForest)
library(dplyr)
library(ggplot2)
library(caret)

hotel_ml_rf <- read.csv("./data/hotel_ml.csv") %>% select(-agent)
```

#### Check NAs
```{r rf_NA, echo = F, message=F}
hotel_ml_rf %>% is.na %>% melt %>% 
  ggplot(data = .,aes(x = Var2, y = Var1)) + geom_tile(aes(fill = value,width=0.95)) +
  scale_fill_manual(values = c("lightblue","white")) + theme_minimal(14) + 
  theme(axis.text.x  = element_text(angle=45, vjust=.75), 
        legend.position='None',
        legend.direction='horizontal',
        panel.grid.major=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        plot.margin=unit(c(.1,.1,.1,.1), "cm")) + 
  labs(title = 'Missing values in each column', subtitle='represented as white bars')
```
From the plot we do not any NAs in the dataset.

#### Random Forest Model
```{r randomForest, echo = F, message = F}
#Data Partition
set.seed(123)
ind <- sample(2, nrow(hotel_ml_rf), replace = T, prob = c(0.8, 0.2))
hotel_ml_rf$is_canceled <- as.factor(hotel_ml_rf$is_canceled)
train <- hotel_ml_rf[ind==1,]
test <- hotel_ml_rf[ind==2,]

#Random Forest
set.seed(222)
rf <- randomForest(is_canceled ~ ., data = train, importance = T)
print(rf)
#attributes(rf)

#confusion matrix
rf$confusion %>% knitr::kable(digits = 3)


#Predication & Confusion Matrix - train data
#p1 <- predict(rf, train)
#confusionMatrix(p1,train$is_canceled)


#Predication & Confusion Matrix - test data
p2 <- predict(rf, test)
confusionMatrix(p2,test$is_canceled)
```
Train:Test = 8:2 \ 
The error rate for the model is 14.98%. The accuracy of the model on test set is 84.93%.

#### Error Rate
```{r rf error rate, echo = F, message = F}

#Error rate of Random Forest
plot(rf, main = "Error Rate")

```
                         
From the plot on error rate for random forest model, we can see that the error rate does not change much after 100.

#### Tune mtry
```{r tune mtry, echo = F, message = F}
#Tune mtry
t <-tuneRF(train[,-14], train[,14],
       stepFactor = 0.5,
       plot = TRUE,
       ntreeTry = 200,
       trace = TRUE,
       improve = 0.05)
t

```
Mtry refers to the number of variables available for splitting at each tree node.\
For tuning Mtry, we set paramter _ntreeTry_ (number of trees used at tuning step) to 200, since from the error rate plot we can see that ER does not change much after 200.\
From the plot we can see that the OOB (Out of Bag Error) is very high when Mtry is 1, and goes down at Mtry equal 3, and reaches the bottom at Mtry equals 6, then bounces back.


#### Tuned RF model
```{r tuned model, echo = F, message = F}
rf_tuned <- randomForest(is_canceled ~ ., data = train,
                         importance = TRUE, mtry=6, ntree= 200)
print(rf_tuned)

# Prediction on test set
confusionMatrix(predict(rf_tuned,test), test$is_canceled) 

```

The accuracy of the tuned model on the test set increase 0.01, not much.

#### Variable Importance
```{r var import, echo = F, message = F}
varImpPlot(rf_tuned, main = "Variable Importance")
```


#### Partial Dependece 
```{r partial depe, echo = F, message = F}
partialPlot(rf_tuned, train, lead_time, "1", main = "Partial Dependece on Lead Time for Class 1")
partialPlot(rf_tuned, train, lead_time, "0", main = "Partial Dependece on Lead Time for Class 0")
```
To show the predictive ability of the variable, we take _LeadTime_ as an example.\
The graph above represents the the partial dependece on _LeadTime_ for Class 1 (canceled) and Class 0 (not canceled). \
If the lead time between booking date and entering/canceled date is greater than 100 days, the model tends to predict the booking will not be canceled.











## Discussion 

##### Potential influence factors of cancellation

As indicated by both the distribution plot and proportion plot, we found there still exists somewhat effects on the cancellation brought by our variables of interest. 

For the assigned room type, room A, G, H was found to have highest cancellation proportion for Resort Hotel; room A was found to have highest cancellation proportion for City Hotel. Taking the consideration that this room type was not chosen by guest themselves compared to the reserved room type which is another variable in the raw data set, we could infer that these room types might be defective in some specific perspectives, which cannot meet the expectations and needs for the majority of guests. However, more information and data are needed to further figure out the relationships and problems with each room type for both hotels.  

For the distribution channel, TA/TO (“Travel Agents” and “Tour Operators”) was found to have highest cancellation proportion among all other booking channels. In our real life, there also exists the fact that many people would like to book hotels through some travel agents or tour operators since these channels may provide some extra discounts compared to the official booking channel of the hotel. Otherwise, when people travel to a new place, with the lack of the local information, some people would also choose to book hotels through the local travel agents or tour operators for convenience. However, the result indicated by our investigation will also raise some issues with these two booking channels. Why the majority people who booked hotels through TA/TO channels canceled their order eventually? We could infer that these two channels might exits some problems with their sales means and operation model. For instance, information asymmetry could be a common issue with many unofficial distribution channels. Some unofficial booking channels try to make the environment and facilities of hotel rooms look nicer and offer extra benefits to mislead guests. When guests arrive at the hotel, they might find that the actual condition is not as same as the information provided by these unofficial channels so that they would cancel the order if the actual condition cannot meet their expectations and needs. However, to further explore the issues of TA/TO channels, which lead to the high cancellation proportion, we still need more information and data.

For the meal type, we found that the cancellation proportion was highest among guests who booked HB meal with comparable size for Resort Hotel; for City hotel, the cancellation proportion is highest among guests who booked BB meal. Therefore, we could infer that these meal types of these two hotels might be defective in some specific perspectives, which cannot fulfill the common expectations and needs of the majority of guest. The food safety issue also could be a consideration for this point. But still, more information and data are needed to further explore the potential problems. 

For the repeated guest, the reason why we are interested in this variable is that we want to see how the repeated-guest effect works on the real problem. From our plots, we found that the cancellation proportion is significantly higher among non-repeated guests, which meets the common expectation for the repeated-guest effect since the repeated guest they might be used to the room, facilities and surrounding environments of the hotel. However, looking at the distribution of the guests for both hotels, the repeated-guest is still a much smaller consumer source compared to the non-repeated guests. Therefore, we might infer that these two hotels are lack of their loyal customer base due to some problems in their management or their hotel infrastructure. However, more information about the hotel are needed such as whether they have membership and their customer reviews.


##### The results of our selected variables indicate a few important things: 
* The total number of special requests added during reservation do have a significant relationship with reservation cancellation. Customers who reserved city hotel with zero special requests are more likely to cancel the reservation, compared to the resort hotel. And it follows our expectations; as the city hotel receives most of the reservation for business travelers, who usually have few special requests, but also more likely to modify their travel schedules.

* The relationship between deposit type used for reservation and cancellation analyzed in this report contradicts with out expectation, as well as counters the intuitive: almost all customers reserved rooms with non-refundable deposit would cancel the reservation for both resort and city hotel. Both the logistic regression and EDA bar plot indicate such relationship. Therefore, we applied cross validation to determine the validity of deposit type, and the result demonstrates that the model with or without deposit type variable shows the same prediction accuracy. As a result, we would not consider deposit type for future analysis. 

* The analysis of lead time of reservation shows that the longer the lead time, the higher probability that customers would cancel the reservation, for both city and resort hotel. This result meets our expectation. However, there are many outliers that the reservations were reserved 300 days before check in, we are not sure about their influence to the result. 

<br>

##### Limitations: 
* We analyzed the data for the entire year 2016, but it is not generalizable to predict for future circumstance, especially during the pandemic and post pandemic period, like 2020 and 2021, at least. During such period, the number of cancellations would significantly increase because of the pandemic, rather than the effects we analyzed above. And customers would ask for more special requests for their safety. All these factors reduces the prediction accuracy of the analysis. 
