---
title: "Project Report"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true

---

```{r setup, include=FALSE, echo=FALSE, message=F}
library(dplyr)
library(tidyverse)
library(plotly)
library(ggplot2)
library(modelr)
library(ROCR)
library(pROC)
library(purrr)
library(readr)
library(tidyr)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Motivation 
## Related work
* Antonio N., Almeida A., Nunes L. (2019). Hotel Booking Demand Datasets. Data in Brief. Vol 22. 41-49.

## Initial Questions
* Where do the guests come from?   
* How does price vary over the year?    
* What are the potential factors which could influence the cancellation of both hotels?

## Data
```{r data,echo=FALSE, message=F}
hotel_whole_df <- read.csv("./data/hotel_bookings.csv",na = c("", "NA", "NULL")) 
hotel_df <- hotel_whole_df %>% filter(arrival_date_year==2016)
```

> **Primary Data Source** 

 [Hotel Booking Demand Dataset](https://www.sciencedirect.com/science/article/pii/S2352340918315191) 
 
> **Variable of Interest**

The original dataset contains information from 2015 to 2017. To avoid the caveat of time series and for a speedier performance of our model, we only included the year 2016 for our analysis. In order to identify any seasonality trend over the year, we created a new variable containing the arrival season of the guests. We then mutated the character variables as factors and additionally some numeric variables that would be better explained as factors. To avoid collinearlity, we kept arrival season and exluced arrival month. Finally, to satisfy our aim of predicting booking cancelation, we selected the following variables that we desired to include in our model:  

  1. Outcome:\
 _IsCanceled_: Values indicating if the booking was canceled (1) or not (0).\
 \
  2. Potential Predictors:\
  * _Categorical Variables_:\
    + _ArrivalDateMonth_: Month of arrival date with 12 categories: “January” to “December”.
    + _CustomerType_: Types of booking, assuming one of the four categories:\
-Contract:  when the booking has an allotment or other type of contract associated with it (long-term stay);\
-Group: when the booking is associated with a group;\
-Transient: when the booking is not part of a group or contract and is not associated with other transient bookings (these are guests who are predominantly on-the-move and seek short (and often urgent) hotel-stays)\
-Transient-party:  when the booking is transient but is associated with at least another transient booking.\
    + _DepositType_: Indication on if the customer made a deposit to guarantee the booking. The variable can assume 3 categories:
-No Deposit;\
-Non Refund: Deposit value equals or exceeds the total stay cost;\
-Refundable: Deposit value under the total stay cost.\
    + _DistributionChannel_: Booking distribution channel (“TA” means travel agent; “TO” means tour operator).\
    + Hotel: City Hotel or Resort Hotel. \
    + _MarketSegment_: Market segment designation (“TA” means travel agent; “TO” means tour operator). \
    + _Meal_: Type of meal booked.\
-Undefined/SC: no meal package;\
-BB: Bed & breakfast;\
-HB: Half board (breakfast and one other meal)\
-FB: Full board (breakfast, lunch, and dinner)\
    + _ReservedRoomType_: Code of room type reserved. Code is presented instead of designation for anonymity reasons.\
    
  * _Numeric Variables_:\
    + _ADR_: Average daily rate.\
    + _BookingChanges_: Number of changes/amendments made to the booking from the moment the booking was entered on the PMS until the moment of check-in or cancellation. \
    + _PreviousCancellations_: Number of previous bookings that were canceled by the customer prior to the current booking.\
    + _LeadTime_: Number of days that elapsed between the entering date of the booking into the PMS and the arrival date. \
    + _TotalOfSpecialRequests_: Number of special requests made by the customer (e.g. twin bed or high floor)\


> **Data Cleaning and Preprocessing**

  + Convert missing value and NAs (both capital and lower-case) to na for R recognition.
  + Convert numeric to factor variables if needed.
  + Create _ArrivalSeason_ variable based on _ArrivalDateMonth_ to investigate the effect of seasonality on hotel booking demands.
  + Create _FalseRoomAssignment_ variable by comparing the difference between _ReservedRoomType_ and _AssignmedRoomType_. 
  




## Exploratory Analysis
We first used a choropleth map and a donut chart to visualize the proportions of guests' nationality. We observed that most guests come from Europe and top three nationality are Portuguese, with British and French guest come second and third.     

We then tried to explore how the ADR(average daily rate) change over time. Therefore, we used group by summary to produce the mean ADR by month, and plotted these values using line chart. To illustrate how the ADR varies, we also added confidence band to the average line. We observed that the ADR is pretty stable across the years for city hotel, which is reasonable according to the booking demand pattern. For resort hotel, we see a sharp increase in ADR from May to November with the price level peaking during summer. This finding is also intuitive since we expect the demand for resort hotel increases in summer. 



## Statistical Analysis 

### Variable Selection
The dataset is quite large (119,390 rows, 32 variables). For a speedy performance of our models, we first selected instances in year 2016 alone (the dataset ranges from 2015 to 2017), which resulted in 56,707 rows.\
\
For categorical variables, we selected 8 out of 14 variables based on our hypothesis on their association with _IsCanceled_.\
\
For numeric variables, we selected 5 variables based on their AUC scores.
```{r AUC,echo=FALSE, message=F }

 hotel_df_xyl <- read.csv("./data/hotel_bookings.csv",na = c("", "NA", "NULL")) %>%
   select(where(is.numeric)) %>% select(-agent)

 variable_list <- colnames(hotel_df_xyl)
 variable_list <- variable_list[-1]

 compute_auc <- function(x) {
   as.numeric(auc(roc(hotel_df_xyl[["is_canceled"]][!is.na(hotel_df_xyl[[x]])], hotel_df_xyl[[x]][!is.na(hotel_df_xyl[[x]])])))
 }

 auc_result<-as_tibble(t(combn(variable_list, 1)),
           .name_repair = ~ c("var_name"))%>%
   mutate(
     auc = map_dbl(.x = var_name, compute_auc)
   ) %>%
   arrange(-auc)
 auc_result %>% knitr::kable(digits = 3)

```

An AUC score equals 0.5 implies that the variable is not discriminable between two classes. Thus, we chose variables that had the greatest distances to 0.5.\
From the AUC score table above, we chose _LeadTime_, _TotalOfSpecialRequests_, _PreviousCancellations_, _adr_, and _BookingChanges_ for our models.



### Logistic Regression  
```{r data import,echo=FALSE, message=F}
hotel_ml = 
  read_csv("./data/hotel_ml.csv") %>% 
  mutate(
    is_canceled = as.factor(is_canceled)
  ) %>% 
  mutate_if(is.character, as.factor) %>% 
  select(-arrival_date_month) %>% 
  drop_na() 
```

```{r, include = FALSE, Fecho=FALSE, message=F}
fullmod = glm(is_canceled ~. , data = hotel_ml, family = binomial())
finalmod = 
  stats::step(fullmod, direction = "both") %>% 
  broom::tidy() %>% 
  knitr::kable()
```
We first tried to predict the output using Logistic regression. The first model we used included all the possible predictors we selected. To perform variable selection, we used stepwise regression. The result shows that the all of our existing variables have statistical significance. We thus fitted this model to our dataset, and generated the following output for the variables with significant odds ratios:  
```{r test/train split,echo=FALSE, message=F, warning=FALSE}
set.seed(1234)
train_df = sample_frac(hotel_ml, 0.8)
test_df = anti_join(hotel_ml, train_df)

logistic_fit =  
  train_df %>% 
  glm(is_canceled ~. , data = ., family = binomial())

logistic_fit %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate)) %>%
  select(term, log_OR = estimate, OR, p.value) %>% 
  filter(OR > 1.9 | OR < 0.2) %>% 
  arrange(desc(OR)) %>% 
  knitr::kable(digits = 3, caption = "Logistic Regression Output") 
```
**Comment** As we can see from the output, deposit type: non refund has the largest odds ratio(173.389), indicating that the guests with non refundable test are significantly more likely to cancel the bookings compared to no deposit guests. This finding is rather counter-intuitive and also raises the question about the validity of the dataset.  
<br>

#### Variable Importance 
We furthur generated the most important variables of our model:  
```{r variable importance,echo=FALSE, message=F}
caret::varImp(logistic_fit) %>%
  arrange(desc(Overall)) %>%
  filter(Overall > 10) %>% 
  knitr::kable(digits = 3, caption = "Variable Importance")
```

##### We will look more into the three most important features:  
* Total special requests
* deposit type 
* lead time
* adr 

<br> 


##### The Special Request Effect

```{r specialrequest, echo=F, message=F, warning=F}
special_request_df <- 
  hotel_df %>%
  mutate(is_canceled = recode(is_canceled, "1" = "Canceled", "0" = "Not Canceled")) %>%
  select(hotel, is_canceled, total_of_special_requests)


special_request_prop <- 
  special_request_df %>%
  group_by(total_of_special_requests, hotel) %>%
  summarize(
    guest_total = n(), 
    guest_canceled = sum(is_canceled == "Canceled")
  ) %>%
  mutate(
    prop_test = map2(.x = guest_canceled, .y = guest_total, ~prop.test(x = .x, n = .y)),
    tidy_test = map(.x = prop_test, ~broom::tidy(.x))
  ) %>%
  select(-prop_test) %>%
  unnest() %>%
  select(total_of_special_requests, hotel, estimate, conf.low, conf.high)


special_request_prop_plot <- 
  special_request_prop %>%
  mutate(text_label = str_c("Number of Special Requests: ", total_of_special_requests, ", Cancellation Proportion: ", estimate)) %>%
    plot_ly(
    x = ~total_of_special_requests, y = ~estimate, color = ~hotel,
    type = "bar", colors = "viridis", text = ~text_label
  ) %>%
  layout(title = "Canceled Proportion with Total Number of Special Request",
         xaxis = list(title = "Number of Special Requests"),
         yaxis = list(title = "Proportion of Cancellation", range = c(0, 1)),
         barmode = "group"
)

special_request_prop_plot
```

Based on the proportion plot, we found that, for city hotel, the cancellation proportion is the highest when the guests didn't ask for any special request. Similarly, for resort hotel, the cancellation proportion is also higher when guests didn't ask for any special request. 
 
<br> 

##### The Deposit Type Effect

```{r deposit, echo=F, message=F, warning=F}
deposit_df <- 
  hotel_df %>%
  mutate(is_canceled = recode(is_canceled, "1" = "Canceled", "0" = "Not Canceled")) %>%
  select(hotel, is_canceled, deposit_type)


deposit_prop <- 
  deposit_df %>%
  group_by(deposit_type, hotel) %>%
  summarize(
    guest_total = n(), 
    guest_canceled = sum(is_canceled == "Canceled")
  ) %>%
  mutate(
    prop_test = map2(.x = guest_canceled, .y = guest_total, ~prop.test(x = .x, n = .y)),
    tidy_test = map(.x = prop_test, ~broom::tidy(.x))
  ) %>%
  select(-prop_test) %>%
  unnest() %>%
  select(deposit_type, hotel, estimate, conf.low, conf.high)


deposit_prop_plot <- 
  deposit_prop %>%
  mutate(text_label = str_c("Deposit Type: ", deposit_type, ", Cancellation Proportion: ", estimate)) %>%
    plot_ly(
    x = ~deposit_type, y = ~estimate, color = ~hotel,
    type = "bar", colors = "viridis", text = ~text_label
  ) %>%
  layout(title = "Canceled Proportion with Deposit Type",
         xaxis = list(title = "Deposit Type"),
         yaxis = list(title = "Proportion of Cancellation", range = c(0, 1)), 
         barmode = "group"
)

deposit_prop_plot
```

Based on the proportion plot, we found that for both resort and city hotels, reservations with non-refundable deposit have about 94% and 99% of cancellation respectively, which is significantly higher than other two types of deposits. For reservations with refundable deposit, we found that the cancellation of city hotel is relatively higher than that of resort hotel. This counter intuitive finding confirmed with the previous odds ratios. We therefore suspected the validity of this variable.  

<br> 

#### The Lead Time Effect

```{r leadtime, echo=F, message=F, warning=F}
leadtime <-   
  hotel_df %>%
  select(hotel, is_canceled, lead_time) %>%
  mutate(is_canceled = recode(is_canceled, "1" = "Canceled", "0" = "Not Canceled"))

leadtime_box <- 
  leadtime %>%
  plot_ly(
    x = ~is_canceled, y = ~lead_time,
    type = "box", color = ~hotel, 
    colors = "viridis"
  ) %>%
  layout(title = "Cancellation Distribution with Lead Time",
         xaxis = list(title = "cancellation"),
         yaxis = list(title = "lead time (days)"),
         boxmode = "group"
         )

leadtime_box
```
Based on the box plot, we found that, for both city and resort hotels, the medians of lead time among being canceled reservations are relatively higher than those of not being canceled. 

<br>

##### The ADR Effect
```{r echo=F, message=F, warning=F}
hotel_df %>% 
  select(is_canceled, adr) %>% 
  group_by(adr) %>% 
  summarize(
    total = n(),
    canceled = sum(is_canceled == 1),
    prop = round(canceled/total, 2)
  ) %>% 
  filter(total > 10) %>% 
  drop_na() %>% 
  mutate(
    fv = lm(prop ~ adr, .) %>%  fitted.values()
  ) %>% 
  mutate(text_label = str_c("Cancellation Proportion: ", prop, ", ADR: ", adr)) %>%
  plot_ly(x = ~adr, y = ~prop, mode = "markers", text = ~text_label) %>%
  add_markers(y = ~prop) %>% 
  add_trace(x = ~adr, y = ~fv, mode = "lines") %>% 
  layout(title = "Average Daily Rate (ADR) and Cancellation", 
         xaxis = list(title = "ADR"), 
         yaxis = list(title = "Cancellation Proportion"),
         showlegend = F)
```

From this plot, there is no clear trend detected. We see that the proportion of cancelation slight increases with increase in ADR.
    
#### Cross Validation  
Since we are questioning the validity of deposit type, we fitted the logistic regression models without the deposit type. We then compared the prediction accuracy of the two models:
```{r ,echo=FALSE, message=F, warning = FALSE}
logistic_fit_2 = 
  train_df %>% 
  glm(is_canceled ~.-deposit_type , data = ., family = binomial())

test_df %>% 
  add_predictions(logistic_fit) %>% 
  mutate(
    prob_model_1 = boot::inv.logit(pred),
    pred_mod1 = 
      case_when(
        prob_model_1 < 0.5 ~ 0,
        prob_model_1 >= 0.5 ~ 1
         )) %>% 
  select(-pred, -prob_model_1) %>% 
  add_predictions(logistic_fit_2) %>% 
   mutate(
    prob_model_2 = boot::inv.logit(pred),
    pred_mod2 = 
      case_when(
        prob_model_2 < 0.5 ~ 0,
        prob_model_2 >= 0.5 ~ 1
         )) %>% 
  select(is_canceled, pred_mod1, pred_mod2) %>% 
  summarise(
    model1_accuracy = sum(is_canceled==pred_mod1)/ nrow(test_df),
    model2_accuracy = sum(is_canceled==pred_mod2)/ nrow(test_df)
  ) %>% 
  knitr::kable(digits = 3, 
               caption = "Model Accuracy",  
               col.names = c("Model1 accuracy", "Model2 accuracy"))
```
We can see that the two models have roughly the same prediction accuracy! We also generated the ROC for comparison:  
```{r echo=FALSE, message=F}
library(ROCR)

test.predicted.m1 <- predict(logistic_fit, newdata = test_df, type = "response")

test.predicted.m2 <- predict(logistic_fit_2, newdata = test_df, type = "response")

par(mfrow=c(1, 2))

prediction(test.predicted.m1, test_df$is_canceled) %>%
  performance(measure = "tpr", x.measure = "fpr") %>%
  plot()

prediction(test.predicted.m2, test_df$is_canceled) %>%
  performance(measure = "tpr", x.measure = "fpr") %>%
  plot()
```

<br>

#### Model Diagnosis 
```{r}
train_df =
  broom::augment(logistic_fit_2) %>% 
  mutate(index = 1:n()) 

ggplot(data = train_df, aes(index, .std.resid, color = "viridis")) + 
  geom_point(alpha = .5) +
  geom_ref_line(h = 3) + 
  labs(title = "Residual Deviance plot")

plot(logistic_fit_2, which = 4, id.n = 5)
```

<br>
We created the residual deviance plot to detect outliers. And as shown in the plot, those standardized residuals that exceed 3 represent possible outliers and may deserve closer attention. Similarly, plotting the cook's distance also signals some potential outliers in our dataset.  



## Discussion 

##### The results of our selected variables indicate a few important things: 
* The total number of special requests added during reservation do have a significant relationship with reservation cancellation. Customers who reserved city hotel with zero special requests are more likely to cancel the reservation, compared to the resort hotel. And it follows our expectations; as the city hotel receives most of the reservation for business travelers, who usually have few special requests, but also more likely to modify their travel schedules.

* The relationship between deposit type used for reservation and cancellation analyzed in this report contradicts with out expectation, as well as counters the intuitive: almost all customers reserved rooms with non-refundable deposit would cancel the reservation for both resort and city hotel. Both the logistic regression and EDA bar plot indicate such relationship. Therefore, we applied cross validation to determine the validity of deposit type, and the result demonstrates that the model with or without deposit type variable shows the same prediction accuracy. As a result, we would not consider deposit type for future analysis. 

* The analysis of lead time of reservation shows that the longer the lead time, the higher probability that customers would cancel the reservation, for both city and resort hotel. This result meets our expectation. However, there are many outliers that the reservations were reserved 300 days before check in, we are not sure about their influence to the result. 

<br>

##### Limitations: 
* We analyzed the data for the entire year 2016, but it is not generalizable to predict for future circumstance, especially during the pandemic and post pandemic period, like 2020 and 2021, at least. During such period, the number of cancellations would significantly increase because of the pandemic, rather than the effects we analyzed above. And customers would ask for more special requests for their safety. All these factors reduces the prediction accuracy of the analysis. 
